Fail Faster

### 2025-03-21 paper/paper.tex
blah

\section{What are Generator Libraries, and why are they Slow?}
\label{section:motiv}
% \jwc{In this section we answer: ``What are monadic generator libraries, and why are they slow?''}
% \bcp{That's actually not a bad section title!}

@@

The runtime of \camlname generators across different input sizes is summarized in \figref{fig:time}.
Our \camlname type-derived BST generator achieves speedups ranging from 1.08--1.29x, which increase
dramatically to 6.13--9.65x when combined with \csplitmix. The insertion-based BST generator sees
1.13--1.21x speedups with staging alone, and 4.32--5.36x when also
using \csplitmix. The single-pass BST generator benefits more significantly from
staging, with speedups of 2.31--2.34x, which rise to 8.61--8.95x when combined with
\csplitmix. We attribute the differential in performance speedup to the fact that
BST strategies rely on relatively few calls to generator combinator functions,
but sample extensively from the \rand; notably, single-pass BST, which uses the largest
number of calls to generator combinators, also sees the greatest benefit from staging.

For STLC, staging accounts for the bulk of performance gains, yielding
 1.71--2.64x speedups for the type-derived STLC generator, and 4.60--5.46x for
 the well-typed generator. When adding \csplitmix, these numbers increase to
 2.60--5.15x and 4.83--7.75x, respectively. This result makes sense because
 STLC generators, particularly the well-typed generator, sample very little from
 randomness libraries, but extensively on generator library functions.

The boolean list generator experiences 3.61--7.10x speedups with
staging, and its performance actually slightly \textit{decreases} when
combined with \csplitmix (3.43--6.83x). This is likely because sampling
booleans is ``cheap'' enough that the overhead of crossing the FFI barrier is
greater than that of generating values directly in OCaml.

>>>>

\section{What are Generator Libraries, and why are they Slow?}
\tr{This needs to introduce PBT WAY better.}
\label{section:motiv}
% \jwc{In this section we answer: ``What are monadic generator libraries, and why are they slow?''}
% \bcp{That's actually not a bad section title!}

@@

The runtime of \camlname generators across different input sizes is summarized in \figref{fig:time}.
Our \camlname type-derived BST generator achieves speedups ranging from $1.08-1.29\times$, which increase
dramatically to $6.13-9.65\times$ when combined with \csplitmix. The insertion-based BST generator sees
$1.13-1.21\times$ speedups with staging alone, and $4.32-5.36\times$ when also
using \csplitmix. The single-pass BST generator benefits more significantly from
staging, with speedups of $2.31-2.34\times$, which rise to $8.61-8.95\times$ when combined with
\csplitmix. We attribute the differential in performance speedup to the fact that
BST strategies rely on relatively few calls to generator combinator functions,
but sample extensively from the \rand; notably, single-pass BST, which uses the largest
number of calls to generator combinators, also sees the greatest benefit from staging.

For STLC, staging accounts for the bulk of performance gains, yielding
 $1.71-2.64\times$ speedups for the type-derived STLC generator, and $4.60-5.46\times$ for
 the well-typed generator. When adding \csplitmix, these numbers increase to
 $2.60-5.15\times$ and $4.83-7.75\times$, respectively. This result makes sense because
 STLC generators, particularly the well-typed generator, sample very little from
 randomness libraries, but extensively on generator library functions.

The boolean list generator experiences $3.61-7.10\times$ speedups with
staging, and its performance actually slightly \textit{decreases} when
combined with \csplitmix ($3.43-6.83\times$). This is likely because sampling
booleans is ``cheap'' enough that the overhead of crossing the FFI barrier is
greater than that of generating values directly in OCaml.
